---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data
```{r}
library(ggplot2)
data <- read.csv('activity.csv')
aggdata <- aggregate(steps~date, data, FUN=sum, na.rm=TRUE)

```


## What is mean total number of steps taken per day?

###First get a histogram of the data to give an idea of where the median is, and then compute explicitly.

```{r}

myplot <-ggplot(aggdata, aes(x=steps))
myplot<- myplot+ geom_histogram(binwidth=1200, col='blue', fill='white')
myplot <- myplot + labs(title="Daily Number of Steps Taken", x="Steps", y="Frequency (Days)")
myplot <- myplot +theme_minimal()
myplot
```

###With my binning choice, the mean and median appear to be in the range [10,000, 11,000].
```{r}
datmean <- mean(aggdata$steps, na.rm=TRUE)
datmed <- median(aggdata$steps, na.rm=TRUE)
datmean
datmed
```

## What is the average daily activity pattern?

###Aggregate the data again, but use FUN=mean
```{r}
aggdata <- aggregate(steps~interval, data, FUN=mean, na.rm=TRUE)
myplot<- ggplot(aggdata, aes(x=interval, y=steps)) +geom_line()
myplot <- myplot +theme_minimal()
myplot <- myplot + labs(title="Daily Number of Steps Taken", x="Time of Day (HHMM)", y="Steps Taken")
myplot
```

###This person on average took the most steps sometime near 8AM.

```{r}
mostactive <- aggdata$interval[which.max(aggdata$steps)]
mostactive
```

## Imputing missing values

### Ignoring NA values introduces some bias into the dataset because the person was likely still somewhat active during those times that the measurement failed.  A simple strategy for inferring where the missing values lie is to determine the median for that time interval.


###Get the number of rows that need to have an NA replaced.  There are 2304 NA values to replace.  

```{r}
summary(data)
```

### Use tapply to replace the NAs

```{r}
mynas <- is.na(data$steps)
avg_interval <- tapply(data$steps, data$interval, median, na.rm=TRUE, simplify=TRUE)
data$steps[mynas] <-avg_interval[as.character(data$interval[mynas])]
aggdata <- aggregate(steps~date, data, FUN=sum, na.rm=TRUE)
newmean <- mean(aggdata$steps)
newmean
newmedian <- median(aggdata$steps)
newmedian
myplot <-ggplot(aggdata, aes(x=steps))
myplot<- myplot+ geom_histogram(binwidth=1200, col='blue', fill='white')
myplot <- myplot + labs(title="Daily Number of Steps Taken", x="Steps", y="Frequency (Days)")
myplot <- myplot +theme_minimal()
myplot

aggdata <- aggregate(steps~interval, data, FUN=mean, na.rm=TRUE)
myplot<- ggplot(aggdata, aes(x=interval, y=steps)) +geom_line()
myplot <- myplot +theme_minimal()
myplot <- myplot + labs(title="Daily Number of Steps Taken", x="Time of Day (HHMM)", y="Steps Taken")
myplot
mostactive <- aggdata$interval[which.max(aggdata$steps)]
mostactive
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
data$day_of_week <- weekdays(as.Date(data$date))
weekend_days <- c("Saturday", "Sunday")
data$day_type <- ifelse(data$day_of_week %in% weekend_days, "Weekend", "Weekday")
weekend_data <- subset(data, data$day_type == "Weekend")
head(weekend_data)
weekday_data <- subset(data, data$day_type == "Weekday")
we_agg <- aggregate(steps~date, weekend_data, FUN=sum, na.rm=TRUE)
wd_agg <- aggregate(steps~date, weekday_data, FUN=sum, na.rm=TRUE)
mean(we_agg$steps)
mean(wd_agg$steps)
we_agg <- aggregate(steps~interval, weekend_data, FUN=mean, na.rm=TRUE)
wd_agg <- aggregate(steps~interval, weekday_data, FUN=mean, na.rm=TRUE)
# Combine data frames
colnames(we_agg)[2] <- "Weekend_Steps"
colnames(wd_agg)[2] <- "Weekday_Steps"
plotframe <- merge(wd_agg, we_agg, by='interval')
#And plot the two time series on the same plot in order to compare
#Activity at different times of the week.  
myplot<- ggplot(plotframe, aes(interval, y=value, color="Time of Week")) 
myplot <- myplot +geom_line(aes(y=Weekday_Steps, col="Weekday Steps"), linetype='dashed')
myplot <- myplot +geom_line(aes(y = Weekend_Steps, col="Weekend Steps"), linetype='solid')
myplot <- myplot + scale_color_manual(values = c('red', 'blue'))
myplot <- myplot + scale_linetype_manual(values = c('dashed','solid'))
myplot <- myplot +theme_classic()
myplot <- myplot + guides(colour=guide_legend(override.aes=list(linetype=c(2,1))))
myplot <- myplot + labs(title="Daily Number of Steps Taken", x="Time of Day (HHMM)", y="Steps Taken", color = "Time ofWeek")

myplot
```